Notes and other text

### 2023-03-31 doc/manuscript/translocation.qmd

Results from frog translocations indicated that adult survival was often relatively high and that nearly all populations showed evidence of successful reproduction (described above).
Although suggestive of population establishment, a decade or more of surveys may be necessary to confirm that populations are in fact self-sustaining [@joseph2018].
To extend our inferences of population establishment beyond those possible from the CMR data, we developed a population viability model.
Specifically, to test whether the observed yearly adult survival probabilities in translocated populations were sufficient to promote long-term viability, we built a stage-structured matrix model that captured known frog demography and included demographic and environmental stochasticity.
We parameterized the model using CMR data from translocated populations and known life history values in this system (@tbl-param_values SI).
Given observed yearly adult survival probabilities (from analysis of CMR data) and a baseline yearly recruitment probability of 0.3 (i.e., the probability that, following survival, juveniles successfully recruit to the adult class), our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 1.15-1.36 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
When we reduced the probability of successful recruitment $\omega$ to 0.1 (a highly conservative estimate), long-run growth rate of five of these six populations was still greater than 1 (@fig-viability A).
In contrast, for the other six translocated populations, where yearly adult survival was substantially less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A).
A sensitivity analysis of long-run growth rate $\lambda$ to recruitment probability $\omega$ and yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$ showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).
<!--# May need to add some text early in this section to clarify why so far in this manuscript we have focursed on yearly survival probability of translocated adults, but the model focuses more on yearly survival probability of naturally recruited adults (e.g., sensitivity analysis, etc.). It is my understanding that you are generally assuming that these two probabilities are equivalent and interchangable, but additional clarification would be helpful. Might also require some clarification in Table 2.-->

Given demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of juvenile recruitment to the adult class $\omega$ (the transition that we expect to be the most subject to environmental stochasticity in the presence of Bd), our model <!--# Is this a different model, now with demographic stochasiticity and environmental variability in juvenile recruitment, or the same model used to estimate lamda in the previous paragraph? Clarify.  -->predicted that, following a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities had 50-year extinction probabilities of less than 1 (@fig-viability B).
Therefore, even under conservative assumptions about recruitment probability $\omega$, these populations had some probability of persistence over 50 years following a single introduction.
<!--# The wording in the previous two sentences seems overly cautious in describing the extinction probability (i.e., "less then 1", "some probability of persistence". In fact, as the next sentence makes clear, for reasonable probabilities of recruitment, the extinction probabilities are much less than 1. Consider tweaking wording to clarify this. I'd argue for a relatively strong initial statement that we then fine tune only slightly, instead of a weak statement that we then strengthen a lot. Initial impressions can be hard to change. --> For these six populations, as $\omega$ increased from 0.1 to 0.3, 50 year extinction probability was less 0.5 <!--# Based on Fig. 5, looks like extinction propability is < 0.5 for all six populations only for omega > ~0.22. Reword as necessary. -->, indicating strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment success (@fig-viability B).
In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history (@fig-viability-supp B; population 70550, surveyed for 16 years).
Overall, our model demonstrates that given observed yearly adult survival probabilities of resistant frogs in non-natal locations, 50% of our translocated populations have a high probability of population growth and long-term viability in the presence of Bd.
Moreover, five of the seven populations translocated after 2013 were predicted to have long-term viability, compared to only one of five populations translocated prior to 2013.
Thus, there is some trend in our improved ability to select appropriate habitat to promote long-term viability of translocated, resistant populations.

@@

Translocations of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species.
We identified donor populations from which adult frogs ($\geq$ 40 mm snout-vent length) could be collected using several years of visual encounter surveys (VES) and skin swab collections [@knapp2016], as well as landscape genetic information [@poorten2017].
Results indicated that these populations contained several hundred *R. sierrae* adults and hundreds or thousands of tadpoles, and that frog abundances were recovering from previous Bd-caused declines [@knapp2016].
As is typical for recovering MYL frog populations, Bd prevalence in the donor populations was high (`r round(bdprev_min, 2)`--`r round(bdprev_max, 2)`) and Bd infection intensity ("load", expressed as median by site: `r round(bdload_min, 2)`--`r round(bdload_max, 2)`) was two or more orders of magnitude below the level at which frog mortality is expected (log~10~(load) $\approx$ 5.78 ITS copies per swab)[@vredenburg2010; @joseph2018].
Recipient sites to which frogs were translocated were chosen based on previous *R. sierrae* presence (determined from VES and/or museum records) or characteristics that suggested high quality habitat for this species [@knapp2005].

During each translocation, adult frogs were collected from the donor population and measured, weighed, swabbed, and PIT tagged.
Frogs were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately once per month during the summer active season (median number of visits per summer = `r round(surveys_yr_median)`; range = `r surveys_yr_min`--`r surveys_yr_max`) and assessed using diurnal CMR surveys and VES.
CMR surveys allowed quantification of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and subadult abundance.
During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location.
During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (median number of secondary periods per primary period = `r round(robust_year_median)`; range = `r robust_year_min`--`r robust_year_max`), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was low, we conducted only a single-day CMR survey.
Within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed during the first capture, and during subsequent captures were only identified and released.
Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods.
For details of survey and swab analysis methods, see @joseph2018.

@@

We estimated survival of translocated frogs, recruitment of new frogs into the adult population, and adult population size using a Bayesian open-population Jolly-Seber mark-recapture model with known additions to the population (i.e., translocated cohorts), as implemented by the mrmr package [@joseph2019] and using R Statistical Software [v4.4.4, @rsoftware2022].
The model tracks the states of *M* individuals that comprise a superpopulation made up of real and pseudo-individuals [see @joseph2018 for details].
The possible states of individuals include "not recruited", "alive", and "dead".
The possible observations of individuals include "detected" and "not detected".
We assume that individuals that are in the "not recruited" or "dead" states are never detected (i.e., there are no mistakes in the individual PIT tag records).
For all models, we used mrmr defaults for priors, number of chains (4), and warmup and post-warmup iterations (2000 for each).
We evaluated convergence of the Markov chain Monte Carlo (MCMC) algorithm using trace plots and Gelman-Rubin statistics (Rhat).
Code used to construct CMR datasets and fit models for each translocated population is available from the following GitHub repository: [https://github.com/SNARL1/cmr-analysis](https://github.com/SNARL1/cmr-analysis.).

@@

To quantify changes in Bd load, we used a workflow similar to that described above (in "Predictors of post-translocation frog survival").
Briefly, the response variable was before-translocation load and the predictor variable was after-translocation load.
We compared models with and without group-level effects, and group-level effects included pit_tag_id, pit_tag_id nested within site_id, or pit_tag_id nested within translocation_id.
All models used a negative binomial family.
Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the following GitHub repository: <https://github.com/SNARL1/translocation>.
<!--# Does it make more sense to only analyze the model with pit_tag_id as the group-level effect. That is the model that is most analogous to a paired t-test. -->

@@

To determine the implications of observed 1-year adult survival on the long-term viability of populations established via translocation, we developed a population model for *R. sierrae*.
Our central question was: how does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations?
We developed a model that tracked seven state variables of an *R. sierrae* population: the density of translocated adults ($A_T$), the density of adults naturally recruited into the population ($A_R$), the density of first year larvae ($L_1$), the density of second year larvae ($L_2$), the density of third year larvae ($L_3$), the density of first year juveniles ($J_1$), and the density of second year juveniles ($J_2$).
We divided adults into two classes $A_T$ and $A_R$ because there is evidence that the survival probability of translocated adults and naturally recruited adults differs ([@fig-compare_surv_probs] : SI).

@@

The parameters in this model are yearly survival probability $\sigma_{\cdot}$, where the "$\cdot$" indicates a particular state variable, the probability that a female frog reproduces in a given year $p_F$, the number of eggs produced by a female frog in a year that successfully hatch $F$, the probability of a first year tadpole remaining as a tadpole $p_{L1}$, the probability of a second year tadpole remaining as a tadpole $p_{L2}$, and the probability of a first year juvenile remaining as a juvenile $p_{J1}$.
The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.

@@

We extracted yearly survival probabilities for translocated adults $\sigma_{A_T}$ and naturally recruited adults $\sigma_{A_R}$ from the CMR model used to estimate population-level parameters of translocated populations (see "Estimation of adult survival and population size" above).
Although we observed yearly variability in adult survival *within* a population, the magnitude of this variability was small compared to *among* population variability ([@fig-translocation-survival]).
Thus, we did not include yearly within-population variability in adult survival in this analysis.
However, within a population there was substantial yearly variability in the successful recruitment of adults, greater than what we would expect from Poisson variability about a mean value.
Therefore, we allowed for yearly variability in the probability of successful recruitment $\omega$ (additional details provided in "Estimating model parameters" below).

@@

The baseline parameter values for the model and how they were estimated are given in @tbl-param_values : SI.
Parameters $\sigma_{A_T}$ and $\sigma_{A_R}$ were extracted directly from our CMR model described previously in this study.
For populations where we had a sufficient number of pit-tagged, naturally-recruited adults, we observed that $\sigma_{A_T}$ and $\sigma_{A_R}$ could be notably different, with $\sigma_{A_R} > \sigma_{A_T}$ ([@fig-compare_surv_probs] : SI).
For populations lacking any naturally-recruited adults, we were unable to directly estimate $\sigma_{A_R}$, and instead set $\sigma_{A_R} = \sigma_{A_T}$.

@@

![For each of the naive and recovering MYL frog populations, pie charts showing allele frequencies for the stringent set of 11 outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, SRFBP1, and (C) RIN3. Charts are superimposed on a map of the Sierra Nevada study area, with Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada shown in light gray. The inset map locates the study area in California. <!--# Allie: Add compass rose and format distance bar to match that in Yosemite map. Relocate pie charts to avoid the perception that all recovering populations are west of naive populations. In legend, provide interpretation of "Homozygous reference", "Homozygous alternate", and "Heterozygous". Try moving inset map to lower left portion of panel - more room there to enlarge inset? How about identifying each population as R. muscosa or R. sierrae using squares and dots? --> <!--# Bree comment: Minor but the little set-in map in the central panel may be too small to be useful. Roland response:  -->](figures/allele_maps.jpg){#fig-allelefrequencies}

![Results from the splined window analyses showing outlier windows for $\pi_{diff}$ (A) and *F~ST~* (B). Outlier windows for either $\pi_{diff}$ or *F~ST~* are shown in blue and outlier windows for both $\pi_{diff}$ and *F~ST~* are shown in red. (C) Magnified Contig19 from A showing two adjacent outlier windows for $\pi_{diff}$ 12.9Mb upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue line). (D) Magnified Contig8 from from B showing the *F~ST~* outlier window that includes the outlier SNPs TCF19 and VARS. This region of the genome contains 10 annotated genes known to occur in the extended MHC Class I and II regions. <!--# As I commented in the Results text, why where the regions highlighted in C and D chosen? -->](figures/spline_manhattan_figure.jpg){#fig-spline-manhattan}

@@

![A population viability analysis of translocated *R. sierrae* that quantifies the long-run growth rate and extinction probability of frog populations: (A) The predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated *R. sierrae*. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (given by the dashed line). (B) The predicted fifty-year extinction probabilities of the twelve translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to A. There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript_v2.jpg){#fig-viability}

![Sensitivity analysis and empirical validation of the stage-structured *R. sierrae* model. (A) The elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$, $\omega$, and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) The 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

@@

![Results from the exome capture selection study. (A) Map of sampling locations in the Sierra Nevada study area showing Bd-naive populations in red and recovering populations in blue. Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) are shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada is shown in light gray. (B) PCA calculated from binary SNPs showing genomic relationship of samples. Numeric labels match those in (A) and are colored as in (A). (C) qqplot showing observed and expected p-values for 148,307 SNPs and indels as calculated in GEMMA. Dashed line shows Bonferroni corrected p-value for threshold of 0.05 (liberal) and 0.01 (stringent) outliers <!--# Add line for 0.05, if that distinction would be useful. -->. (D) Manhattan plot showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by genomic position and contigs are sorted by size. Red line shows Bonferroni corrected p-value for threshold of 0.01. <!--# Show line for threshold of 0.05 also? -->Outlier SNPs above this threshold are highlighted and labeled. Bold labels indicate at least one non-synonymous SNP is present in that gene. <!--# Map (panel A) is duplicated here and in the mapped allele frequency figure. I think it can be removed from this figure. Capitalize panel labels. Label/identify both the liberal and stringent outlier sets? -->](figures/map_qqplot_manhattan.jpg){#fig-selectionresults}

![Map showing the locations of translocated and donor populations in Yosemite National Park (park boundary indicated by black polygon). Symbol shapes indicate the donor population used for each translocation site. To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of Yosemite within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation). <!--# Make sure all points are inside YOSE polygon -->](figures/map_translocation_points.png){#fig-yosemap width="450"}

![For the best model describing the average 1-year survival of frogs in the first translocation to a site as a predictor of survival of frogs in all subsequent translocations to that site, estimated posterior density curves and shaded 95% uncertainty intervals for the intercept and single predictor variable. A variable is considered an important predictor if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m2b.png){#fig-transsurvival-postdens width="450"}

![For frogs translocated to each of the 12 recipient sites, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation.](figures/bdload_beforeafter.png){#fig-bdload-beforeafter width="450"}

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for the intercept and all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens width="450"}

>>>>

Results from frog translocations indicated that adult survival was often relatively high and that nearly all populations showed evidence of successful reproduction (described above).
Although suggestive of population establishment, a decade or more of surveys may be necessary to confirm that populations are in fact self-sustaining [@joseph2018].
To extend our inferences of population establishment beyond those possible from the CMR data, we developed a population viability model.
Specifically, to test whether the observed yearly adult survival probabilities in translocated populations were sufficient to promote long-term viability, we built a stage-structured matrix model that captured known frog demography and included demographic and environmental stochasticity.
We parameterized the model using CMR data from translocated populations and known life history values in this system (@tbl-param_values SI).
Given observed yearly adult survival probabilities (from analysis of CMR data) and a baseline yearly recruitment probability of 0.3 (i.e., the probability that, following survival, juveniles successfully recruit to the adult class), our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 1.15-1.36 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
When we reduced the probability of successful recruitment $\omega$ to 0.1 (a highly conservative estimate), long-run growth rate of five of these six populations was still greater than 1 (@fig-viability A).
In contrast, for the other six translocated populations, where yearly adult survival was substantially less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A).
A sensitivity analysis of long-run growth rate $\lambda$ to recruitment probability $\omega$ and yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$ showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).
<!--# May need to add some text early in this section to clarify why so far in this manuscript we have focused on yearly survival probability of translocated adults, but the model focuses more on yearly survival probability of naturally recruited adults (e.g., in the sensitivity analysis, etc.). It is my understanding that you are generally assuming that these two probabilities are equivalent and interchangable, but additional clarification would be helpful. Might also require some clarification in Table 2. You describe this in more detail in the Methods, but it seems that some additional clarification might be helpful here. -->

Given demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of juvenile recruitment to the adult class $\omega$ (the transition that we expect to be the most subject to environmental stochasticity in the presence of Bd), our model <!--# Is this a different model, now with demographic stochasiticity and environmental variability in juvenile recruitment, or the same model used to estimate lamda in the previous paragraph? Clarify. I know from the Methods that you developed different versions of the model, some deterministic and some stochastic, but some clarification might be useful here too. -->predicted that, following a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities had 50-year extinction probabilities of less than 1 (@fig-viability B).
Therefore, even under conservative assumptions about recruitment probability $\omega$, these populations had some probability of persistence over 50 years following a single introduction.
<!--# The wording in the previous two sentences seems overly cautious in describing the extinction probability (i.e., "less than 1", "some probability of persistence". In fact, as the next sentence makes clear, for reasonable probabilities of recruitment, the extinction probabilities are substantially less than 1. Consider tweaking wording to clarify this. I'd argue for a relatively strong initial statement that we then fine tune only slightly, instead of a weak statement that we then strengthen a lot. Initial impressions can be hard to change. --> For these six populations, as $\omega$ increased from 0.1 to 0.3, 50 year extinction probability was less 0.5 <!--# Based on Fig. 5, looks like extinction propability is < 0.5 for all six populations only for omega > ~0.22. Reword as necessary. -->, indicating strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment success (@fig-viability B).
In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history (@fig-viability-supp B; population 70550, surveyed for 16 years).
Overall, our model demonstrates that given observed yearly adult survival probabilities of resistant frogs in non-natal locations, 50% of our translocated populations have a high probability of population growth and long-term viability in the presence of Bd.
Moreover, five of the seven populations translocated after 2013 were predicted to have long-term viability, compared to only one of five populations translocated prior to 2013.
Thus, there is some trend in our improved ability to select appropriate habitat to promote long-term viability of translocated, resistant populations.

@@

Translocations of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species (@fig-yosemap).
We identified donor populations from which adult frogs ($\geq$ 40 mm snout-vent length) could be collected using several years of visual encounter surveys (VES) and skin swab collections [@knapp2016], as well as landscape genetic information [@poorten2017].
Results indicated that these populations contained several hundred *R. sierrae* adults and hundreds or thousands of tadpoles, and that frog abundances were recovering from previous Bd-caused declines [@knapp2016].
As is typical for recovering MYL frog populations, Bd prevalence in the donor populations was high (`r round(bdprev_min, 2)`--`r round(bdprev_max, 2)`) and Bd infection intensity ("load", expressed as median log~10~(load) by site: `r round(bdload_min, 2)`--`r round(bdload_max, 2)`) was two or more orders of magnitude below the level at which frog mortality is expected (log~10~(load) $\approx$ 5.78 ITS copies per swab)[@vredenburg2010; @joseph2018].
Recipient sites to which frogs were translocated were chosen based on previous *R. sierrae* presence (determined from VES and/or museum records) or characteristics that suggested high quality habitat for this species [@knapp2005].

During each translocation, adult frogs were collected from the donor population and measured, weighed, swabbed, and PIT tagged.
Frogs were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately once per month during the summer active season (median number of visits per summer = `r round(surveys_yr_median)`; range = `r surveys_yr_min`--`r surveys_yr_max`) and assessed using diurnal CMR surveys and VES.
CMR surveys allowed quantification of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and subadult abundance.
During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location.
During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (median number of secondary periods per primary period = `r round(robust_year_median)`; range = `r robust_year_min`--`r robust_year_max`), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed during the first capture, and during subsequent captures were only identified and released.
Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods.
For details of survey and swab analysis methods, see @joseph2018.

@@

We estimated survival of translocated frogs, recruitment of new frogs into the adult population, and adult population size using a Bayesian open-population Jolly-Seber mark-recapture model with known additions to the population (i.e., translocated cohorts), as implemented by the mrmr package [@joseph2019] and using R Statistical Software [v4.4.4, @rsoftware2022].
The model tracks the states of *M* individuals that comprise a superpopulation made up of real and pseudo-individuals [see @joseph2018 for details].
The possible states of individuals include "not recruited", "alive", and "dead".
The possible observations of individuals include "detected" and "not detected".
We assume that individuals that are in the "not recruited" or "dead" states are never detected (i.e., there are no mistakes in the individual PIT tag records).
For all models, we used mrmr defaults for priors, number of chains (4), and warmup and post-warmup iterations (2000 for each).
We evaluated convergence of the Markov chain Monte Carlo (MCMC) algorithm using trace plots and Gelman-Rubin statistics (Rhat).
Code used to create CMR datasets and fit models for each translocated population is available from the following GitHub repository: [https://github.com/SNARL1/cmr-analysis](https://github.com/SNARL1/cmr-analysis.).

@@

To quantify changes in Bd load, we used an analysis workflow similar to that described above.
Briefly, the response variable was before-translocation load and the predictor variable was after-translocation load.
We compared models with and without group-level effects, and group-level effects included pit_tag_id, pit_tag_id nested within site_id, or pit_tag_id nested within translocation_id.
All models used a negative binomial family.
Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the following GitHub repository: <https://github.com/SNARL1/translocation>.
<!--# Does it make more sense to only analyze the model with pit_tag_id as the group-level effect. That is the model that is most analogous to a paired t-test. -->

@@

To determine the implications of observed 1-year adult survival on the long-term viability of populations established via translocation, we developed a population model for MYL frogs.
Our central question was: How does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations?
We developed a model that tracked seven state variables of a frog population: density of translocated adults ($A_T$), density of adults naturally recruited into the population ($A_R$), density of first-year tadpoles ($L_1$), density of second-year tadpoles ($L_2$), density of third-year tadpoles ($L_3$), density of first-year juveniles ($J_1$), and density of second-year juveniles ($J_2$).
We divided adults into two classes $A_T$ and $A_R$ because there is evidence that the survival probability of translocated adults and naturally recruited adults differs ([@fig-compare_surv_probs] : SI).

@@

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ ("$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year $p_F$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile $p_{J1}$.
The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.
<!--# Can the omega parameter be added to the list of other parameters: "..., probability of a first-year juvenile remaining as a juvenile pj1, and probability of a juvenile successfully recruiting to an adult omega." Listing omega separately seems unnecessary, at least to me. -->

@@

We extracted yearly survival probabilities for translocated adults $\sigma_{A_T}$ and naturally recruited adults $\sigma_{A_R}$ from the CMR model used to estimate population-level parameters of translocated populations (see "Estimation of adult survival and population size" above).
Although we observed yearly variability in adult survival *within* a population, the magnitude of this variability was small compared to *among* population variability ([@fig-translocation-survival]).
Thus, we did not include yearly within-population variability in adult survival in this analysis.
However, within a population there was substantial yearly variability in the successful recruitment of adults, greater than what we would expect from Poisson variability around a mean value.
Therefore, we allowed for yearly variability in the probability of successful recruitment $\omega$ (additional details provided in "Estimating model parameters" below).

@@

The baseline parameter values for the model and how they were estimated are given in @tbl-param_values : SI.
Parameters $\sigma_{A_T}$ and $\sigma_{A_R}$ were extracted directly from our CMR model described previously in this study.
For populations where we had a sufficient number of PIT-tagged, naturally-recruited adults, we observed that $\sigma_{A_T}$ and $\sigma_{A_R}$ could be notably different, with $\sigma_{A_R} > \sigma_{A_T}$ ([@fig-compare_surv_probs] : SI).
For populations lacking sufficient numbers of naturally-recruited adults, we were unable to directly estimate $\sigma_{A_R}$, and instead set $\sigma_{A_R} = \sigma_{A_T}$.

@@

![For each of the naive and recovering MYL frog populations, pie charts showing allele frequencies for the stringent set of 11 outlier SNPs from 7 distinct genes: (A) LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, SRFBP1, and (C) RIN3. Charts are superimposed on a map of the Sierra Nevada study area, with Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada shown in light gray. The inset map locates the study area in California. <!--# Allie: Add compass rose and format distance bar to match that in Yosemite map. Relocate pie charts to avoid the perception that all recovering populations are west of naive populations. In legend, provide interpretation of "Homozygous reference", "Homozygous alternate", and "Heterozygous". How about identifying each population as R. muscosa or R. sierrae using squares and dots, respectively? The current version of the figure ignores the species identities, but including them might help with interpretation of results. Try moving inset map to lower left portion of panel - more room there to enlarge inset?  Remove distance scale from inset map. --> <!--# Bree comment: Minor but the little set-in map in the central panel may be too small to be useful. Roland response:  -->](figures/allele_maps.jpg){#fig-allelefrequencies}

![Results from the splined window analyses showing outlier windows for $\pi_{diff}$ (A) and *F~ST~* (B). Outlier windows for either $\pi_{diff}$ or *F~ST~* are shown in blue and outlier windows for both $\pi_{diff}$ and *F~ST~* are shown in red. (C) Magnified Contig19 from (A) showing two adjacent outlier windows for $\pi_{diff}$ 12.9Mb upstream of the RIN3 outlier SNP (indicated with a dashed vertical blue line). (D) Magnified Contig8 from from (B) showing the *F~ST~* outlier window that includes the outlier SNPs TCF19 and VARS. This region of the genome contains 10 annotated genes known to occur in the extended MHC Class I and II regions. <!--# As I commented in the Results text, numerous outlier regions were identified, so how and why were the regions highlighted in C and D chosen? Legend text that describes the outlier categories needs additional editing to clarify the categories. My attempt feels only partially successful. Capitalize plot identifiers (A, B, C, D). -->](figures/spline_manhattan_figure.jpg){#fig-spline-manhattan}

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated frogs. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (given by the dashed line). (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505. <!--# Remove period after plot labels, e.g., "A", not "A.". -->](figures/pop_viability_figures_for_manuscript_v2.jpg){#fig-viability}

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$, $\omega$, and *F (*number of eggs produced by a female frog in a year that successfully hatch) <!--# Necessary to define F (when sigma and omega are not defined)? Maybe instead just cite "Table 2 SI"?) -->. Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

<!--# Remove period after plot labels -->

@@

![Results from the exome capture selection study. (A) Map of sampling locations in the Sierra Nevada study area showing Bd-naive populations in red and recovering populations in blue. Yosemite, Kings Canyon, and Sequoia National Parks (from north to south) are shown in dark gray, and the range boundary for MYL frogs in this portion of the Sierra Nevada is shown in light gray. (B) PCA calculated from binary SNPs showing genomic relationship of samples. Numeric labels and colors match those in (A). (C) qqplot showing observed and expected p-values for 148,307 SNPs and indels as calculated in GEMMA. Dashed line shows Bonferroni corrected p-value for threshold of 0.05 (liberal) and 0.01 (stringent) outliers <!--# Add line for 0.05, if that distinction would be useful. -->. (D) Manhattan plot showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by genomic position and contigs are sorted by size. Red line shows Bonferroni corrected p-value for threshold of 0.01. <!--# Show line for threshold of 0.05 also? -->Outlier SNPs above this threshold are highlighted and labeled. Bold labels indicate at least one non-synonymous SNP is present in that gene. <!--# Map (panel A) is duplicated here and in the mapped allele frequency figure. I think it can be removed from this figure. Capitalize panel labels. Label/identify both the liberal and stringent outlier sets? -->](figures/map_qqplot_manhattan.jpg){#fig-selectionresults}

![Map showing the locations of translocated and donor MYL frog populations in Yosemite National Park (park boundary indicated by black polygon). Symbol shapes indicate the donor population used for each translocation site. To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of Yosemite within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation). <!--# Make sure all points are inside YOSE polygon -->](figures/map_translocation_points.png){#fig-yosemap width="450"}

![For the best model describing the average 1-year survival of frogs in the first translocation to a site as a predictor of survival of frogs in all subsequent translocations to that site, estimated posterior density curves and shaded 95% uncertainty intervals for the intercept and single predictor variable. A variable is considered an important predictor if the associated uncertainty interval does not overlap zero (indicated by the dashed red line).](figures/mcmc_areas_m2b.png){#fig-transsurvival-postdens width="450"}

![For frogs translocated to each of the 12 recipient sites, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation. Box plot details are as in @fig-cond-effects. ](figures/bdload_beforeafter.png){#fig-bdload-beforeafter width="450"}

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for the intercept and all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (indicated by the dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens width="450"}

![Comparison of the average yearly adult survival probabilities for adults translocated to each of three sites versus adults that were naturally recruited at each site (as a result of reproduction by translocated frogs). In contrast to @fig-translocation-survival, these are not survival probabilities from the first year following translocation, but instead represent averaged survival probabilities across multiple years. Points are median estimates and error bars give the 90% uncertainty intervals around the estimates. All estimates are derived from the mrmr package. <!--# Use 95% uncertainty intervals for consistency? Also, use dark blue instead of gray for naturally-recruited? Might show up better. -->](figures/compare_surv_probs.jpg){#fig-compare_surv_probs}
